### 2016-04-19 paper/sections/CaseStudy.tex
paper structure

\section{Case Study}

Talk about the 3 case studies

Talk about Pretty Printer case study last and discuss issues with 
extensible transformations.