### 2016-04-19 paper/sections/RelatedWork.tex

\section{Related Work}

\paragraph{Deep and Shallow Embeddings}

\paragraph{Modularity of DSLs?}
Data types a la Carte; Finally Tagless; 
Object Algebras

\paragraph{Family Polymorphism}

\paragraph{ThisType}

\paragraph{Multiple Inheritance}