### 2016-04-19 paper/sections/Tool.tex

\section{Code Generation}

How to generate code for family polymorphism!