### 2016-04-19 paper/sections/Discussion.tex
paper structure

\section{Discussion and Limitations}

- Transformations and Binary methods (but argue that 
these types of operations are not used in shallow DSLs 
any way!).

- Java syntax not great for DSLs. 
operator overloding; infix operators (but this is a Java limitation; 
not a limitation of our approach).

- Limitations of the tool: separate compiltions and ... 
generics not fully implemented.

- How to apply our approach to other OO language 
  - C\# has annotations, but can you do AST rewriting?
  - Scala has macros;
  - Other languages?