### 2016-04-19 paper/sections/Tool.tex
Delegating tasks

\section{Code Generation}

Haoyuan should write this part.

How to generate code for family polymorphism!