### 2016-04-23 paper/sections/Overview.tex
Some text for S3

\section{Overview}

\begin{comment}
Weixin writes this one.

Go over jeremy's examples, maybe having only 3 diagram 
constructs instead of 5 for space reasons.

Use five constructs to show extensibility.

Think of how to introduce our tool? Using the Jeremy's examples? 
or introducing before with some other examples?

\end{comment}

%As language evolves, the need of new syntax often arises.
DSLs may evolve along the time with new syntax and semantics introduced.
This is essentially the Expression Problem (EP)~\cite{wadler}, which requires
the host language equipped with two dimensions of extensibility.
To illustrate, we start with a subset of the DSL presented in Gibbons and Wu's
paper with a single semantic function. And then we show how to extend this
language with full language constructs and additional semantics both in Haskell
and in Java. The key difference is that the extension on semantics for Java is
\emph{modular} whereas the Haskell approach needs to modify existing code.

\subsection{A DSL for Parallel Prefix Circuits}
Initially, the language contains three language constructs: a primitive
\emph{fan} and two combinators \emph{beside} and \emph{stretch}. Their types are
given as follows:

\lstinputlisting[language=haskell,linerange=2-5]{./code/shallowCircuit.hs}%APPLY:CTYPE
\emph{fan n} contains \emph{n} vertical wires with the first wire connected to
all the remaining wires from top to bottom; \emph{beside c1 c2} joins two circuits
horizontally; \emph{stretch ns c} insertes additional wires into the circuit
\emph{c} by summing up \emph{ns}.
For example, a circuit constructed with these three operator \emph{stretch [2,2,2] (fan 3) `beside` fan 1} is visualized in Figure~\ref{}.
The semantics of the language is given below:

\lstinputlisting[language=haskell,linerange=9-12]{./code/shallowCircuit.hs}%APPLY:CIRCUIT1
which calculates the width of a circuit, i.e. the number of vertical wires in a circuit.
The corresponding Java implementation is shown in Figure~\ref{}.

\lstinputlisting[linerange=9-9]{../src/paper/Circuit.java}%APPLY:CIRCUIT
\lstinputlisting[linerange=13-16]{../src/paper/Circuit.java}%APPLY:FAN
\lstinputlisting[linerange=20-26]{../src/paper/Circuit.java}%APPLY:STRETCH
\lstinputlisting[linerange=30-36]{../src/paper/Circuit.java}%APPLY:BESIDE
The translation is straightforward. The only difference is that classes are
replaced with interfaces and fields become unimplemented getter methods. This
change is vital for extensibility, which will be discussed later in this section.

\subsection{Adding New Syntax}
Shallow embedding makes it easy to \emph{modularly} add new syntax both in FP and OOP.
Suppose that we would like to add a primitive \emph{identity n} which contains
\emph{n} isolated wires and a combinator \emph{above c1 c2} which combines two
circuits of the same width vertically.

In Haskell, we can define two more cases for the semantic function \texttt{identity} and
\texttt{above} like this:
\lstinputlisting[linerange=51-52]{./code/shallowCircuit.hs}%APPLY:SYNTAX_HS
It is also trivial in Java. We just need to define two new interfaces
\texttt{Identity} and \texttt{Above} that both extend \texttt{Circuit} and
implement the \texttt{width} method:

\lstinputlisting[linerange=40-43]{../src/paper/Circuit.java}%APPLY:ID
\lstinputlisting[linerange=47-51]{../src/paper/Circuit.java}%APPLY:ABOVE

\subsection{Adding New Semantics}
Adding new semantics, however, becomes hard.
For example, we would like to have an operation that checks
whether the circuit is of correct size. We say a circuit is \emph{well-sized}
only if: 1) for \emph{fan n}, \emph{n} is a positive number; 2) for
\emph{beside c2}, both \emph{c1} and \emph{c2} are well-sized; 3) for
\emph{stretch ns c} \emph{c} is well-sized and its width is the same as the
length of \emph{ns}.

Gibbons and Wu solved this problem similar to how we add \texttt{peval} to
\texttt{Exp}, where they used a pair instead of a record to capture the new
semantic domain, then defined the two semantic functions simultaneously for each
case and split the definitions through projections on the tuple in the end.
This solution, however, modifies existing code, breaking the requirement of EP.

Conversely, the support of covariant type-refinements and inheritance for OO
languages allows us to add new semantics \emph{modularly}:
%Different from records or tuples, interfaces are extensible.
\lstinputlisting[linerange=56-58]{../src/paper/Circuit.java}%APPLY:CIRCUITWS
\lstinputlisting[linerange=62-64]{../src/paper/Circuit.java}%APPLY:FANWS
\lstinputlisting[linerange=68-73]{../src/paper/Circuit.java}%APPLY:STRETCHWS
\lstinputlisting[linerange=77-83]{../src/paper/Circuit.java}%APPLY:BESIDEWS

Interface \texttt{CircuitWS} extends the original interface and declare a new
semantic function \texttt{wellSized} inside.
Then all existing cases should extend this new interface with the semantic
function implemented. Note that multiple inheritance is used for composing
existing semantic functions and also for methods whose return type is
\texttt{Circuit} should be refined to \texttt{CircuitWS} so that we can call
\texttt{wellSized} method on such fields.

extensibility.
for return types. This is \ref{EPtrivially}
(that is why we implement language constructs as interfaces rather than classes)

\subsection{\name's Support}
Although manually . It has some drawbacks.
It is cubersome to .
To retain extensibility, all should be kept as interfaces, which means that
additional classes should be created for each of them so that objects can be constructed.
Once new semantics is introduced, all the
Programmers would not get warned may forget to implement
Also, it is cubersome to express the dependencies boilerplate and do
type-refinements manually.

\name addresses all these problems through code instrument.
By using \name, we can rewrite the extensions in the way shown in Figure\ref{}.

Figure~\ref{} shows the \name's implementation of Circuit with extensions.
Note that \name generates \texttt{of} method for class-like interfaces for
constructing objects. Inside \texttt{of}, an anonymous class is created with all
field-like methods implemented and its instance is returned.
Moreover, \name has a good support family polymorphism.
Client users need only explicitly declared the dependencies on other families
through (...). \name will re-declare all members from the base families in the extended family.
And the dependencies between members of families are automatically deducted and
implicitly expressed in member re-declaration.