### 2016-04-26 paper/sections/Tool.tex
Text. Code generation.

\section{Code Generation}

%Haoyuan should write this part.
%How to generate code for family polymorphism!

In this section, we present an overview of the code that \name generates. The syntax and type system are
consistent with the Java language. We use translation functions to illustrate our code generation.
To make it clearer, we split the process of code generation into two parts, in which case we introduce a
new annotation \lstinline{@FamilyOf}, which is not defined in the source code but only helps to explain
the translation. In our implementation, the annotation processing is a combination of \name
and \lstinline{@FamilyOf}. Below are the translation functions:

\begin{lstlisting}
  (*$\llbracket$*)@Obj interface (*$I_0$*) extends (*$\overline{I_s}$*) {(*$\overline{meth}$*) (*$\overline{I}$*)}(*$\rrbracket$*)
= @ObjOf interface (*$I_0$*) extends (*$\overline{I_s}$*) {(*$\overline{meth'}$*) (*$\overline{I'}$*)}
\end{lstlisting}

\subsection{@Obj}
\lstinline{@Obj} builds the structure of family polymorphism for the annotated type. More specifically, \lstinline{@Obj} tackles two
tasks: (1) building the dependencies (subtyping relations) between new family members and old ones; (2) refining field types. To this
aim, \lstinline{@Obj} re-declares all member types and field methods from the base families. The detection runs recursively throughout
the nested interfaces in base families. For member types with same names, the re-declaration in the annotated type is generated or modified
with supertypes filled in for dependencies.

\subsubsection{sameMembers}
The auxiliary function \lstinline{sameMembers} generates the declaration of common member types from base families in the annotated type. It
firstly checks if these members are homogeneous in their own hierarchies, and then fuses the old declarations into a new re-declaration, where
all field types are automatically refined. Meanwhile, the code generation fills in the ``extends'' for the re-declaration, to build the inheritance
relationship to the sources.
\begin{itemize}
\item $I_0\in$ \lstinline{sameMembers(}$\overline{I}$\lstinline{)}
    \begin{itemize}
    \item \lstinline{n = name(}$I_0$\lstinline{)}
    \item $\overline{sup}$\lstinline{ = consistent(n, }$\overline{I}$\lstinline{)}
    \item $I_0$\lstinline{ = fuse(n, }$\overline{sup}$\lstinline{, }$\overline{I}$\lstinline{)}
    \end{itemize}
\item $\overline{sup}$\lstinline{ = consistent(n, }$\overline{I}$\lstinline{)}
    \begin{itemize}
    \item \lstinline{I.n = interface n extends }$\overline{sup}$ \{...\} %style
    \item $\forall$\lstinline{I.n}
    \end{itemize}
\item $I_0$\lstinline{ = fuse(n, }$\overline{sup}$\lstinline{, }$\overline{I}$\lstinline{)}
    \begin{itemize}
    \item $I_0$\lstinline{ = interface n extends }$\overline{sup}$ $\{\overline{meth}\ \ \overline{I_1}\}$ %style
    \item $\overline{meth}$\lstinline{ = fuse(}$\overline{I.\overline{meth}}$\lstinline{)}
    \item $\overline{I_1}$\lstinline{ = fuse(}$\overline{I.\overline{I_1}}$\lstinline{)}
    \end{itemize}
\end{itemize}

\subsubsection{$\overline{I'}$}
\begin{itemize}
\item $\forall I_0\in$ \lstinline{sameMembers(}$\overline{I_s}$\lstinline{)}
    \begin{itemize}
    \item If $\exists I\in\overline{I}$, \lstinline{name(}$I_0$\lstinline{) = name(}$I$\lstinline{)}, then \lstinline{fuse2(}$I_0,I$\lstinline{)}$\in\overline{I'}$
    \item Otherwise $I_0\in\overline{I'}$
    \item More: generating \lstinline{extends}
    \end{itemize}
\end{itemize}

\subsubsection{$\overline{meth'}$}
\begin{itemize}
\item $\overline{meth'}$ = $\overline{meth}$\lstinline{ + copyFields}
\end{itemize}

\subsection{@ObjOf}
The second part of annotation processing is abstracted here using the newly introduced annotation
\lstinline{@ObjOf}. It recursively generates constructer methods \lstinline{of} for the annotated interface
and all nested interfaces. Furthermore, \lstinline{@ObjOf} supports getters, void and fluent setters, withers and
the general \lstinline{with} methods. That is to say, \lstinline{@ObjOf} generates a static \lstinline{of} method
for a type, returning an instance of that type. Such an instance is implemented using an anonymous class with auto-generated
implementations for all the methods it supports (stated above). On the other hand, users are expected to put underscores as the
prefix of field methods, and consequently \lstinline{of} identifies these field methods and takes them as its arguments.

\begin{itemize}
\item Recursively
\item Use ``\_'' to distinguish fields and non-field methods
\item Getters, setters, withers, etc
\item \lstinline{of()} method
\end{itemize}