### 2016-05-14 paper/sections/Discussion.tex
More text.

\section{Discussion and Limitations}

Haoyuan writes this one.

- Transformations and Binary methods (but argue that
these types of operations are not used in shallow DSLs
any way!).

- Java syntax not great for DSLs.
operator overloding; infix operators (but this is a Java limitation;
not a limitation of our approach).

- Limitations of the tool: separate compiltions and ...
generics not fully implemented.

- How to apply our approach to other OO language
  - C\# has annotations, but can you do AST rewriting?
  - Scala has macros;
  - Other languages?

\paragraph{Transformations and binary methods} Some other patterns, however, are not captured by our approach.

\paragraph{Limitations} Our tool has certain limitations at a few aspects. As Lombok only provides experimental support
for separate compilation, our \textsf{@Family} does not support separate compilation at this stage, so all the related
interfaces should be included in a single Java file. As for generics, we provide comparatively nice support for generic interfaces
and methods, however, only method typing of non-generic methods is captured, and that of generic ones is delegated to the Java compiler.

On the other hand, it would be nice to have operator overloading and infix operators in Java syntax, so that the code could be
written in a more concise and elegant way. Nevertheless, this is limited by the programming language, and we expect our approach to be oriented to
other OO languages.