### 2016-05-23 paper/sections/CaseStudyPPrinter.tex
Case study: PPrinter.

\subsection{A Prettier Printer}

This case study refactors the Haskell code from \textcolor{red}{(Haoyuan: ref, a prettier printer)}
which implements a functional pretty printer library with high efficiency. In the original code, two
data structures are defined for documents. Here they are defined using shallow DSLs, and packaged into
two base families with \textsf{@Family} annotation applied. These two families \textsf{Family\_Doc} and
\textsf{Family\_Document} correspond to data types \textsf{Doc} and \textsf{DOC} respectively in original code.

\begin{lstlisting}
@Family interface Family_Doc {
	interface Doc {}
	interface Nil extends Doc {}
	interface Text extends Doc {
		String s(); Doc d();
	}
	interface Line extends Doc { int i(); Doc d(); }
}

@Family interface Family_Document {
	interface Document {}
	interface DNil extends Document {}
	interface DConcat extends Document {
		Document d1(); Document d2();
	}
	interface DNest extends Document {
		int i(); Document d();
	}
	interface DText extends Document { String s(); }
	interface DLine extends Document {}
	interface DUnion extends Document {
		Document d1(); Document d2();
	}
}
\end{lstlisting}

With the help of family polymorphism and \textsf{@Family}, we can easily add new operations and integrate them
in child families. Below is an example of encoding two original functions \textsf{layout} and \textsf{fits} in
the shallow embeddings. Note that with the annotation processing, verbose code for building inheritance relations is
automatically generated.
\begin{lstlisting}
@Family interface Family_Doc_LayoutFits extends Family_Doc {
	interface Doc {
		String _layout(); boolean _fits(int w);
	}
	interface Nil {
		default String _layout() { return ""; }
		default boolean _fits(int w) {
			return w >= 0;
		}
	}
	...
\end{lstlisting}
Furthermore, some functions like \textsf{be} and \textsf{beaux} need refactoring before they are encoded in shallow embeddings.
Finally we refactored the pretty printer library using a set of extensible families with shallow DSLs and operations, and our
\textsf{@Family} annotation. However there is a special operation called \textsf{group}, which works like a transformation:
\begin{lstlisting}
@Family interface Family_Document_Flatten extends Family_Document {
	interface Document {
		Document _flatten();
		default Document _group() {
			return DUnion.of(this._flatten(), this);
		}
	}
	...
}
\end{lstlisting}
The method \textsf{\_group} calls the constructor method from \textsf{DUnion}, whereas such a member type will be updated in the
child families, together with its constructors. To ensure type safety we are unable to reuse the code but just copy it and paste
to the child families, and hence code duplication is introduced. The matter of extensible transformations will be further discussed
in the next section.