### 2016-05-31 paper/sections/CaseStudy.tex
Some text on case study

\section{Case Study}

Weixin and Haoyuan.

Talk about the 3 case studies

Talk about Pretty Printer case study last and discuss issues with
extensible transformations.

\subsection{Scans}
% statistics, pictures, implementation
Section~\ref{} is based on our first case study, hence we discuss it briefly here.
In the case study, we strictly follows the flow of the paper and implement all kinds of
interpretations modularly, including \texttt{depth} and \texttt{layout} that are
not discussed in Section~\ref{}.

Starting from a data structure that describes the five constructs without any interpretations, we gradually add
\texttt{depth}, \texttt{width}, \texttt{wellSized}, \texttt{layout} and
\texttt{tlayout} through extending the family.

To make our DSL more user-friendly, we defined some wrappers for constructing
circuits conveniently. Afterwards, we can construct a circuit this way:
\begin{lstlisting}
Circuit circuit =
  fan(2).beside(fan(2))
  .above(fan(2).stretch(2,2))
  .above(id(1).beside(fan(2)).beside(id(1)));
\end{lstlisting}
For primitives like \texttt{fan}, we define a static method and for binary
combinators like \texttt{above}, we define a default method on the \texttt{Circuit}.

Additionally, we extend the language with a \texttt{draw} method which renders a
circuit using Java Swing.
Calling \texttt{circuit.draw()} on the \texttt{circuit} we just defined will
display what Figure~\ref{} shows.

% show Java swing output
\begin{figure}
\end{figure}

\input{sections/CaseStudyPPrinter.tex}

\subsection{Diagrams}
\emph{Diagrams}~\cite{} is a powerful, flexible, declarative Haskell DSL for creating vector graphics.
This case study implements a simplified version of \emph{Diagrams} from
Gibbons's lecture notes ~\cite{}.
The DSL consists of three simpler sub-languages for describing shapes, colors, styles.
% Haskell code
\begin{lstlisting}[language=haskell]
data Shape
  = Rectangle Double Double
  | Ellipse Double Double
  | Triangle Double

type StyleSheet = [Styling]
data Styling
  = FillColour Col
  | StrokeColour Col
  | StrokeWidth Double

data Col = Red | Blue | Bisque | Black | Green | Yellow | Brown

data Picture
  = Place StyleSheet Shape
  | Above Picture Picture
  | Beside Picture Picture
\end{lstlisting}

\paragraph{Shapes}
There are three primitive shapes - rectangle, ellipse and triangle.

\begin{lstlisting}
interface Shape {}
interface Rectangle extends Shape {
    double _x(); double _y();
}
interface Ellipse extends Shape {
    double _rx(); double _ry();
}
interface Triangle extends Shape {
    double _l();
}
\end{lstlisting}

\paragraph{Styles}
A shape can be decorated with drawing styles. A \texttt{StyleSheet} contains a
list of \texttt{Stylings} where fill color, stroke color or stroke width can be specified.
\begin{lstlisting}
interface StyleSheet {
    List<? extends Styling> _stylings();
}
interface Styling {}
interface FillColor extends Styling {
    Col _color();
}
interface StrokeColor extends Styling {
    Col _color();
}
interface StrokeWidth extends Styling {
    double _width();
}
\end{lstlisting}
\texttt{Col} is also a data structure that describes the supported colors.
\begin{lstlisting}
interface Col {}
interface Red extends Col {}
interface Blue extends Col {}
interface Green extends Col {}
interface Yellow extends Col {}
interface Bisque extends Col {}
interface Black extends Col {}
\end{lstlisting}

\paragraph{Pictures}
\begin{lstlisting}
interface Picture {}
interface Place extends Picture {
    Shape _s(); StyleSheet _ss();
}
interface Above extends Picture {
    Picture _p1(); Picture _p2();
}
interface Beside extends Picture {
    Picture _p1(); Picture _p2();
}
\end{lstlisting}

Provided with a \texttt{StyleSheet}, a shape can be lifted to a picture using
\texttt{Place}. And pictures can be merged using \texttt{Above} or
\texttt{Beside}: the former combines two pictures vertically whereas the latter
combines two pictures horizontally.

With these three main data structure, we can draw a picture.

\subsubsection{Rendering}
We render our fancy picture using scalable vector graphics (SVG) as our backend.
To do so, we need to flatten the picture which is recursively constructed.

SVG is based on XML and hence we define to model a XML file.

\paragraph{Transformations}

\subsubsection{Extensions}
\cite{} contains exercises about extensions on the language, e.g.
adding new shapes. The deep embedding approach in Haskell requires modifications on the ADTs
whereas our approach can modularly introduce new language constructs.
.
Each sub-language retains independent extensibility.
For example, we can add a \texttt{show} method for the purpose of debugging.