### 2016-06-17 paper/paper.tex
more text

\documentclass[10pt,preprint,nocopyrightspace]{sigplanconf}

% The following \documentclass options may be useful:

% preprint      Remove this option only once the paper is in final form.
% 10pt          To set in 10-point type instead of 9-point.
% 11pt          To set in 11-point type instead of 9-point.
% numbers       To obtain numeric citation style instead of author/year.

\usepackage{amsmath}
\usepackage{color}
\usepackage{listings}
\usepackage{balance}
\usepackage{amsmath}
\usepackage{stmaryrd}
\usepackage{graphicx}
\usepackage{amssymb}
\usepackage{fancyvrb}
\usepackage{url}
\usepackage{pstricks,pst-node,pst-tree}
\usepackage{theorem}
\usepackage{bbm}
\usepackage{pgf}
\usepackage{multirow}
\usepackage{rotating}
\usepackage{verbatim}
\usepackage{graphicx}
\usepackage{wrapfig}
\usepackage{xspace}
\usepackage{mdwlist}
\usepackage{morefloats}
\usepackage[T1]{fontenc}
\usepackage[scaled=0.85]{beramono}
\usepackage{mathptmx}
\usepackage[normalem]{ulem}
\usepackage{booktabs}
\usepackage{tablefootnote}
\usepackage{enumitem}
\usepackage{stmaryrd}

\newcommand{\name}{\textbf{@Family}\xspace}
\newcommand{\dsl}{\emph{Scans}\xspace}
\newcommand{\interp}{\textsc{Interpreter}\xspace}

%\newcommand{\authornote}[3]{}
\newcommand{\authornote}[3]{{\color{#2} {\sc #1}:#3}}
\newcommand{\weixin}[1]{\authornote{weixin}{cyan}{#1}}
% \setlist[itemize]{noitemsep,nolistsep}
% \setlist[enumerate]{noitemsep,nolistsep}

\lstdefinelanguage{JavaScala}{
  morekeywords={public,int,interface,implements,default,
    abstract,case,void,catch,class,def,static,%
    do,else,extends,false,final,finally,%
    for,if,implicit,import,match,mixin,%
    new,null,object,override,package,%
    private,protected,requires,return,sealed,%
    super,this,throw,trait,true,try,%
    type,var,val,while,yield,with},
  otherkeywords={=>,<-,<\%,<:,>:,\#,@},
  sensitive=true,
  morecomment=[l]{//},
  morecomment=[n]{/*}{*/},
  morestring=[b]",
  morestring=[b]',
  morestring=[b]"""
}

\lstset{ %
language=JavaScala,                % choose the language of the code
columns=fullflexible,
keepspaces=true,
lineskip=-1pt,
basicstyle=\ttfamily\small,       % the size of the fonts that are used for the code
numberstyle=\ttfamily\tiny,      % the size of the fonts that are used for the line-numbers
stepnumber=1,                   % the step between two line-numbers. If it's 1 each line will be numbered
numbers=none,
numbersep=5pt,                  % how far the line-numbers are from the code
backgroundcolor=\color{white},  % choose the background color. You must add \usepackage{color}
showspaces=false,               % show spaces adding particular underscores
showstringspaces=false,         % underline spaces within strings
showtabs=false,                 % show tabs within strings adding particular underscores
%  frame=single,                   % adds a frame around the code
tabsize=2,                  % sets default tabsize to 2 spaces
captionpos=none,                   % sets the caption-position to bottom
breaklines=true,                % sets automatic line breaking
breakatwhitespace=false,        % sets if automatic breaks should only happen at whitespace
title=\lstname,                 % show the filename of files included with \lstinputlisting; also try caption instead of title
escapeinside={(*}{*)},          % if you want to add a comment within your code
keywordstyle=\ttfamily\bfseries,
aboveskip=3pt,
belowskip=-1pt
% commentstyle=\color{Gray},
% stringstyle=\color{Green}
}
\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\conferenceinfo{CONF 'yy}{Month d--d, 20yy, City, ST, Country}
\copyrightyear{20yy}
\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}
\copyrightdoi{nnnnnnn.nnnnnnn}

% Uncomment the publication rights you want to use.
%\publicationrights{transferred}
%\publicationrights{licensed}     % this is the default
%\publicationrights{author-pays}

\titlebanner{banner above paper title}        % These are ignored unless
\preprintfooter{short description of paper}   % 'preprint' option specified.

\title{Shallow EDSLs and Object-Oriented Programming}

\authorinfo{}
           {}
           {}
%\authorinfo{Name}
%           {Affiliation2/3}
%           {Email2/3}

\maketitle

\begin{abstract}

Shallow Embedded Domain Specific Languages (EDSLs) use
\emph{procedural abstraction} to directly encode a Domain Specific
Language into an existing host language. Procedural abstraction has
been argued to be the essence of Object-Oriented Programming (OOP). Given
that Object-Oriented Languages have evolved over more than 50 years
to improve the use of procedural abstraction, they ought to have some
advantages to encode shallow EDSLs.

This paper argues that OOP abstractions, including \emph{inheritance}
and \emph{subtyping}, increase the modularity and reuse of shallow
DSLs when compared to classical procedural abstraction. We make this
argument by taking a recent paper by Gibbons and Wu, where procedural
abstraction is used in Haskell to model a simple EDSL, and we recode
that EDSL in Java. Although from the \emph{syntactical} point of view
there are obvious disadvantages in the Java version, from the semantic
and modularity point of view the Java version has clear advantages.
%To alleviate some of the syntactical disadvantages of Java, we create
%an annotation inspired by \emph{family polymorphism} and a recent
%solution to the Expression Problem.
%The annotation uses transparent code generation techniques to
%automatically eliminate large portions of boilerplate code.
%To further illustrate the applicability of our tool and techniques, we conduct
%several case studies using larger DSLs from the literature.
\end{abstract}

\begin{comment}
\category{CR-number}{subcategory}{third-level}

% general terms are not compulsory anymore,
% you may leave them out
\terms
term1, term2

\keywords
keyword1, keyword2
\end{comment}

%===============================================================================
\input{sections/Introduction.tex}
\input{sections/ShallowOO.tex}
\input{sections/Overview.tex}
\input{sections/Tool.tex}
\input{sections/CaseStudy.tex}
\input{sections/RelatedWork.tex}
\input{sections/Discussion.tex}
%\acks

%Acknowledgments, if needed.

% We recommend abbrvnat bibliography style.

\bibliographystyle{abbrvnat}
\bibliography{paper}

% The bibliography should be embedded for final submission.

\end{document}