### 2016-06-19 paper/sections/Modularity.tex
Draft Section 4

\section{Modular EDSLs in OOP}
This section shows that procedural abstraction mechanisms provided by OO
languages improves the modularity and reusability of EDSLs, in particular
subtyping and inheritance although most OO languages blend them together in one
syntactic form.

\subsection{Subtyping}
Subtyping allows us separate interfaces and implementations, which
encapsulates details of implementations and yields flexible design of EDSLs.
Also, it forms the foundation of the \textsc{Composite} pattern,
which enables us to composite instances of subtypes where a supertype is needed,
following the Liskov substitution principle.

Figure~\ref{code:base} illustrates the use of subtyping,
where all the classes are subtype of the trait \lstinline{Circuit}.

We can easily introduce new language constructs to \dsl by adding a new class to
the hierarchy.
For example, an \lstinline{above} combinator that combines two circuits vertically:
\begin{lstlisting}
class Above(val c1: Circuit, val c2: Circuit) extends Circuit {
  def width = c1.width
}
\end{lstlisting}
The new class works seamlessly with existing classes, as it is also a subtype of
\lstinline{Circuit}. We can construct a circuit by a combination of old
and new constructs:\\
\lstinline{new Beside(new Above(new Fan(2),new Id(2)),new Id(3))}

% type refinements
% record subtyping.

\subsection{Inheritance}
Inheritance is critical for code reuse, which allows
a class to obtain functionalities from other existing classes
without duplicating the code.
Moreover, the inherited functionality can be refined through method overriding
and new functionalities can be introduced via defining new methods and
fields.

We have already shown how to reuse an EDSL through inheritance in Figure~\ref{code:desugar}.
Instead of copying and pasting \lstinline{width} definition case by case
from the old implementation to the new one, we define the new language through inheriting existing
classes, e.g. \lstinline{class Fan extends base.Fan}.
Additionally, we introduce a new method \lstinline{desugar} to the hierarchy.
and gives a default implementation in \lstinline{Circuit} and override it when needed.
Note that we do covariant type-refinements on fields of type \lstinline{Circuit} so that
we can call \lstinline{desugar} method on them.

% Overriding in FP
One may argue that similar code reuse can be simulated in FP through reusing a record instance.
In fact, it is more like composition than inheritance from the OO perspective.
Inheritance plus dynamic dispatch maximize the code reuse for composition.
Consider a simple example:

\begin{lstlisting}[language=haskell]
data T = T { f :: Int, g :: Int }
a = T { f = g a, g = 0 }
b = a { g = 1 }
\end{lstlisting}
Here, we want to reuse \lstinline{a} but ``override'' its \lstinline{g} field when defining \lstinline{b}.
However, the expected result of calling \lstinline{f b} is 1 but it is indeed 0.
The reason is that the \lstinline{f} field ``inherited'' from \lstinline{a} calls
\lstinline{g} on the old instance rather than the new one.

To make it clear, Figure~\ref{code:compare} give the corresponding OO implementation using composition
and inheritance respectively.
\begin{figure}
\begin{tabular}{ll}
\begin{minipage}{.25\textwidth}
\begin{comment}
\begin{lstlisting}
trait T {
  def f: Int
  def g: Int
}
class A extends T {
  def f = g
  def g = 0
}
\end{lstlisting}
\end{comment}
\begin{lstlisting}
object Composition {
  class B(a: A) extends T {
    def f = a.f
    def g = 1
  }
  new B(new A).f // 0
}
\end{lstlisting}
\end{minipage}
&
\begin{minipage}{.25\textwidth}
\begin{lstlisting}
object Inheritance {
  class B extends A {
    override def g = 1
  }
  (new B).f // 1
}
\end{lstlisting}
\end{minipage}
\end{tabular}
\caption{Composition vs inheritance}
\label{code:compare}
\end{figure}
In the case of inheritance, calling \lstinline{f} will execute the overridden \lstinline{g}
as \lstinline{this} points to \lstinline{B} rather than \lstinline{A}.
For compositions, a method can not be reused if when the implementation of what
they depend on changes like \lstinline{f} in \lstinline{A}, even if the logic holds.
OO languages provide these two options for reusing code.

\subsection{Extensibility of EDSLs}
As EDSLs evolve along the time, the need for new syntax and new semantics arises.
This requires host languages equipped with extensibility mechanisms.
OO languages give simple solution to the Expression Problem~\cite{eptrivially16}
and hence are suitable host languages in terms of extensibility.

We have already shown how to extend \dsl on these two dimensions modularly in an
OO language: a new language construct \lstinline{Above} and a new interpretation function \lstinline{desugar}.

Gibbons and Wu introduce new interpretations in a way similar to
Figure~\ref{code:haskell} by appending the definition to a tuple,
which requires revision on the original code.
Though they also present a modular solution based on~\ref{swierstra2008data},
the complexity of encoding increases significantly. Moreover, how to dependent
interpretation modularly is not clear.

\subsection{Discussion}
There are certain kind of methods that are not reusable using the OO approach we
presented, including transformations (or producer methods) and binary
methods as discussed in~\ref{eptrivially16}.
For transformations, we have to refine their return type in extensions and hence
the original code can not be reused, resulting code duplications.
For binary methods like equality, there is no way to refine the argument type.
It is still possible to make it reusable by using virtual types in Scala to capture the type of \lstinline{this}~\ref{zenger}.
However, this may significantly complicate the encoding.