### 2016-06-19 paper/sections/RelatedWorkShort.tex

\section{Related Work}

\paragraph{Deep and Shallow Embeddings}
Shallow embedding yields flexible and concise EDSLs while deep embedding makes
it easy to define optimizations.
There are a lot of work~\cite{svenningsson2012combining,
  Jovanovic:2014:YCD:2658761.2658771, scherr2014implicit} trying to blend these two
approaches to enjoy benefits from both by using the shallow embedding as the
surface language and generate deep embedding version to allow optimizations to
be defined.
%Hofer and Ostermann~\cite{hofer2010modular} propose to provide both embedding through implementing internal and external visitor at the same time so that clients can choose for a particular interpretation;
The OO approach we present retains the simplicity of shallow embedding while
makes it possible to directly implement optimizations.

\paragraph{Modularity of EDSLs}
Languages that solves the EP are capable of being host languages for
implementing modular EDSLs. But many of these solutions require sophisticated
type parameterization, advanced features or heavy encoding which impose
complexity to the DSL implementers or even clients and may be too complicated to use in practice.
For example, solutions based on the \textsc{Visitor}
pattern~\cite{oliveira09modular,hofer2010modular} %more references
Relatively simple solution like Object Algebras~\cite{bruno12oa} may not support
optimizations well.
Our approach, however, is simple, intuitive and supports transformations.