### 2016-06-28 paper/sections/Conclusion.tex
Conclusion

\section{Conclusion}
This paper shows how OO programming improves the modularity of shallow EDSLs.
With the procedural abstraction mechanisms provided by OO languages, various types of
interpretations can be defined modularly. We also show that defining
transformations in shallow embeddings is possible.

There are still certain interpretations that can not be modeled modularly using
our approach, including transformations and binary methods.
For transformations, we have to refine their return type in extensions and
supply a new implementation by duplicating the original code.
For binary methods like equality, there is no way to refine the argument type.
Though with advanced type system features of Scala, these can still be defined
modularly~\ref{zenger05independentlyextensible}. However, the encoding would
become complicated.
Also, we can further eliminate some boilerplate through code generation.
For example, similar inheritance relationships are repeated in each constructs
for extensions.