### 2017-02-11 ICFP2017/Main.tex
Switch to ICFP format

\documentclass[acmlarge, anonymous, review]{acmart}\settopmatter{printfolios=true}

\usepackage{xspace}
\usepackage{url}
\usepackage{syntax}
\newcommand{\dsl}{\textsc{Scans}\xspace}
\newcommand{\interp}{\textsc{Interpreter}\xspace}
\newcommand{\an}[3]{{\color{#2} {\sc #1}:#3}}
\newcommand{\weixin}[1]{\an{weixin}{cyan}{#1}}

%% for comments
\input{Preamble}

\makeatletter\if@ACM@journal\makeatother
%% Journal information (used by PACMPL format)
%% Supplied to authors by publisher for camera-ready submission
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{1}
\acmArticle{1}
\acmYear{2017}
\acmMonth{1}
\acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\else\makeatother
%% Conference information (used by SIGPLAN proceedings format)
%% Supplied to authors by publisher for camera-ready submission
\acmConference[PL'17]{ACM SIGPLAN Conference on Programming Languages}{January 01--03, 2017}{New York, NY, USA}
\acmYear{2017}
\acmISBN{978-x-xxxx-xxxx-x/YY/MM}
\acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}
\fi

%% Copyright information
%% Supplied to authors (based on authors' rights management selection;
%% see authors.acm.org) by publisher for camera-ready submission
\setcopyright{none}             %% For review submission
%\setcopyright{acmcopyright}
%\setcopyright{acmlicensed}
%\setcopyright{rightsretained}
%\copyrightyear{2017}           %% If different from \acmYear

%% Bibliography style
\bibliographystyle{ACM-Reference-Format}
%% Citation style
%% Note: author/year citations are required for papers published as an
%% issue of PACMPL.
\citestyle{acmauthoryear}   %% For author/year citations

% Document starts
\begin{document}

% Title portion
\title{Shallow EDSLs and Object-Oriented Programming}
\subtitle{Functional Pearl}
\author{First1 Last1}
\authornote{with author1 note}          %% \authornote is optional;
                                        %% can be repeated if necessary
\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional
\affiliation{
  \position{Position1}
  \department{Department1}              %% \department is recommended
  \institution{Institution1}            %% \institution is required
  \streetaddress{Street1 Address1}
  \city{City1}
  \state{State1}
  \postcode{Post-Code1}
  \country{Country1}
}
\email{first1.last1@inst1.edu}          %% \email is recommended

\begin{abstract}
Shallow Embedded Domain Specific Languages (EDSLs) use
\emph{procedural abstraction} to directly encode a DSL into an existing host language. Procedural abstraction has
been argued to be the essence of Object-Oriented Programming (OOP).
%Given
%that OO languages have evolved over more than 50 years
%to improve the use of procedural abstraction, they ought to have some
%advantages to encode shallow EDSLs.
This paper argues that OOP abstractions
(including \emph{inheritance}, \emph{subtyping} and
\emph{type-refinement})
increase the modularity and reuse of shallow
EDSLs when compared to classical procedural abstraction. We make this
argument by taking a recent paper by Gibbons and Wu, where procedural
abstraction is used in Haskell to model a simple shallow EDSL, and we recode
that EDSL in Scala. From the \emph{semantic}
and \emph{modularity} point of view the Scala version has clear advantages
over the Haskell version.
\end{abstract}

\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011008</concept_id>
<concept_desc>Software and its engineering~General programming languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10003456.10003457.10003521.10003525</concept_id>
<concept_desc>Social and professional topics~History of programming languages</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~General programming languages}
\ccsdesc[300]{Social and professional topics~History of programming languages}

\keywords{keyword1, keyword2, keyword3}

\maketitle

\input{sections/Introduction}

\input{sections/ShallowOO}

\input{sections/Interpretations}

\input{sections/Casestudy}

% Bibliography
\bibliography{Main}

% History dates
%\received{February 2007}{March 2009}{June 2009}

\end{document}