### 2017-02-20 paper/sections/RelatedWorkShort.tex

\section{Related Work}

\paragraph{Deep and Shallow Embeddings}